Point-process based representation learning for Electronic Health Records

### 2023-02-02 Manuscript_v1.tex
commit

\usepackage{multirow}

@@

@@

\subsection{Event Encoder}
We use a similar transformer architecture [thp] for encoding events. The
advantage of the attention mechanism is that it discards recurrent architecture.
it is necessary to include temporal information. Similar to the original
positional encoding [vaswani], we use a temporal encoding procedure:

@@

\begin{table}[h!]
\centering
\caption{Effect}
\label{tab:my-table}
\begin{tabular}{lllll}
        &        & shp+mark &     & baseline \\ \cline{3-5} 
Dataset & Metric & concat   & sum &          \\ \hline
SO      & NLL    &          &     &          \\
        & AUROC  &          &     &          \\
        & time   &          &     &          \\ \hline
Synthea & NLL    &          &     &          \\
        & AUROC  &          &     &          \\
        & time   &          &     &          \\ \hline
ReTweet & NLL    &          &     &          \\
        & AUROC  &          &     &          \\
        & time   &          &     &          \\ \hline
\end{tabular}
\end{table}

@@

\begin{table}[h!]
\centering
\caption{NLL estimation}
\label{tab:my-table}
\begin{tabular}{lllll}
\hline
        &         &    & Model  &          \\ \cline{3-5} 
Dataset & setting & TE & TE+DAM & TE+noise \\ \hline
P12     & sc      &    &        &          \\
        & mc1     &    &        &          \\
        & mc2     &    &        &          \\ \hline
p19     & sc      &    &        &          \\
        & mc1     &    &        &          \\
        & mc2     &    &        &          \\ \hline
\end{tabular}
\end{table}

@@

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table*}
\centering
\caption{Down stream task}
\label{tab:my-table}
\begin{tabular}{cccccccccccc}
\hline
                     &                      &        &  & F1   &        &  & AUPRC &        &  & AUROC &        \\ \cline{5-6} \cline{8-9} \cline{11-12} 
Dataset              & Setting              & Center &  & DAM  & TE+DAM &  & DAM   & TE+DAM &  & DAM   & TE+DAM \\ \hline
\multirow{8}{*}{P12} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
\multirow{8}{*}{P19} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
\end{tabular}
\end{table*}

>>>>

\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{rotating}
\usepackage{multirow}

@@

\begin{figure*}[!t]
\centerline{\includegraphics{images/model_arch.png}}
\caption{Magnetization as a function of applied field.
It is good practice to explain the significance of the figure in the caption.}
\label{fig1}
\end{figure*}

The entire data consists of $N$ samples  $\mathcal{D}=\{\mathcal{O}_{i}=(\mathcal{E}_{i},\mathcal{S}_{i}  )\}_{i=1}^N$ where $\mathcal{E}_{i}$ and $\mathcal{S}_{i}$ represents event and state sequences, respectively.

Each event sequnce consist of $L$ tuples $\mathcal{E}_{i}=\{ (t_j, e_j)
\}_{j=1}^{L_i}$ where $t_j\in \mathbb{R}  $ is the event timestamp and $ e_j $ is the event mark.

@@

General schematic of TEDAM is depicted in FIG which consists of two separate module: Transformer Event Encoder (TEE) and Deep Attention Module (DAM). TEE encodes timestamp and mark of events using a transformer architecture, while DAM encodes all available information using one-dimensional attention mechanism. Finally, the learned representations are concatenated to be optimized for different loss functions.

\subsection{Event Encoder}
We use a similar transformer architecture [thp] for encoding events with minor modifications. A transformer architecture is capable of .

let . that can be embedded using a trainable embedding matrix.

In the first step, we embed all event marks $E_{emb}=E \times U$ where $E_i \in [L,K] $ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{K \times d_{emb} }  $ is the trainable embedding matrix. In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions. Finally, the input of the transformer encoder will be $x_{emb}=[y(k_j), z(t_j)]$.

Here, we use the standard transformer encoder similar to [vaswani] with masking matrix to prevent the model from looking into the future. we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information before occurence of $j-th$ event.

\subsection{bin of Event Encoder}

it is necessary to include temporal information. Similar to the original
positional encoding [vaswani], we use a temporal encoding procedure:

@@

% Table generated by Excel2LaTeX from sheet 'time concat'
\begin{table}[h!]
    \centering
    \caption{Add caption}
      \begin{tabular}{ccccc}
      \toprule
            &       & \multicolumn{2}{c}{shp+mark} &  \\
  \cmidrule{3-4}    Dataset & Metric & concat & sum   & baseline \\
      \midrule
      SO    & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
      \midrule
      Synthea & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
      \midrule
      ReTweet & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
      \bottomrule
      \end{tabular}%
    \label{tab:addlabel}%
  \end{table}%

@@

% Table generated by Excel2LaTeX from sheet 'state enc'
\begin{table}[h!]
    \centering
    \caption{Add caption}
      \begin{tabular}{ccccc}
      \toprule
            &       & \multicolumn{3}{c}{Model} \\
  \cmidrule{3-5}    Dataset & setting & TE    & TE+DAM & TE+noise \\
      \midrule
      \multirow{3}[2]{*}{P12} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
      \midrule
      \multirow{3}[2]{*}{p19} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
      \bottomrule
      \end{tabular}%
    \label{tab:addlabel}%
  \end{table}%

@@

% Table generated by Excel2LaTeX from sheet 'Sheet3'
\begin{table*}[htbp]
    \centering
    \caption{Add caption}
      \begin{tabular}{ccccccccccc}
      \toprule
            &       &       & \multicolumn{2}{c}{F1} &       & \multicolumn{2}{c}{AUPRC} &       & \multicolumn{2}{c}{AUROC} \\
  \cmidrule{4-5}\cmidrule{7-8}\cmidrule{10-11}    Dataset & Setting & Center & DAM   & TE+DAM &       & DAM   & TE+DAM &       & DAM   & TE+DAM \\
      \midrule
      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
      \midrule
      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
      \bottomrule
      \end{tabular}%
    \label{tab:addlabel}%
  \end{table*}%
